---
title: Diffusion and Contagion
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
library(tidyverse)
```

## Dad Joke

Interviewer: "Where do you see yourself in 5 years?"

<span class='fragment'>
Me: "Listening. I would say listening is my biggest weakness."
</span>

## Housekeeping

> - Self-assessment reflections due
>	- In-person feedback; sign up for spot on my calendar by end of next week
> - Wisdom or Madness of Crowds simulation + R on Thursday

## Upcoming Assignments

> - Exam
>	- Submit two multiple choice or short answer questions by next week
> - Visualization Challenge due next week

## Example questions
> - What is a "multiplex" (or multidimensional) network?
>   - A network where there are multiple types of nodes
>   - A network with high centralization
>   - A network with a high clustering coefficient
>   - A network with multiple types of edges
> - Give an example of how social capital can be exchanged for economic capital.

## Plan for today

> - Diffusion and contagion review
> - Discussion questions
> - R Review

## Review Questions

> - What are some things that are diffused through a social network?
> - What is the difference between simple contagion and complex contagion?

## Simple Contagion

<div class="r-stack">

<div class='fragment'>
```{r, message=F}
create_notable('zachary') |>
  mutate(index = 1:nrow(.N())) |>
  mutate(infected = index %in% c(2,3)) |>
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

<div class='fragment'>
```{r, message=F}
create_notable('zachary') |>
  mutate(index = 1:nrow(.N())) |>
  mutate(infected = index %in% c(2,3)) |>
  mutate(neigh_2 = index %in% neighbors(.G(), 2),
         neigh_3 = index %in% neighbors(.G(), 3)) |>
  mutate(infected = infected | neigh_2 | neigh_3) |>
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

</div>


## Complex Contagion

<div class="r-stack">

<div class='fragment'>
```{r, message=F}
create_notable('zachary') |>
  mutate(index = 1:nrow(.N())) |>
  mutate(infected = index %in% c(2,3)) |>
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

<div class='fragment'>
```{r, message=F}
create_notable('zachary') |>
  mutate(index = 1:nrow(.N())) |>
  mutate(infected = index %in% c(2,3)) |>
  mutate(neigh_2 = index %in% neighbors(.G(), 2),
         neigh_3 = index %in% neighbors(.G(), 3)) |>
  mutate(infected = infected | (neigh_2 & neigh_3)) |>
  ggraph() +
  geom_edge_fan(width = .1) + 
  geom_node_point(aes(color = infected), size = 3) +
  theme_graph()
```
</div>

</div>

## Review Questions

> - What is an example of simple contagion?
> - What is an example of something that spreads via complex contagion?
> - What's an example of when you've been influenced?
> - How does tie strength relate to contagion?
> - What was the argument of the "Special People" chapter?

## R Review


## Discussion Questions

> - If there was a world where we didn't HAVE to have special people or things, would there be a "best" or "most popular" of everything? Or would we simply just have certain paintings, people, and places that are better than others but not necessarily labeled as such?
> - If what we do is influenced by what others do, how can we stay secure in our individuality in social networks?
> - Why is it still so difficult to anticipate what will become popular or go viral when we have so much data now?

## Discussion Questions

> - Have marketers used the findings of the Special People chapter? If not, why not?
> - How can we avoid being led astray by our networks?

## Discussion Questions

> - What do we mean by influencers?
> - What do we mean by post hoc explanations?
> - How can we design better decision-making systems, in government, business, or media, that limit the errors caused by overreliance on "common sense" explanations?
> - Big data and predictability

## Wisdom / Madness of Crowds

> - How does the majority illusion work?
> - How is this related to the spiral of silence?
> - Why is it tough for beliefs to spread across bridges?
> - What is the role of people who aren't easily influenced in contagions?