---
title: Week 12 - Collective Behavior
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
set.seed(0121)
library(igraph)
library(tidygraph)
library(ggraph)
library(tidyverse)
library(networkdata)
# Set graph theme
#df = read_csv('./resources/wisdom_of_crowds_activity.csv') |>
#  filter(is.na(Guess)==F)

BOOK_1_LENGTH = 294
BOOK_2_LENGTH = 374

```

```{r}
df = read_csv('https://docs.google.com/spreadsheets/d/e/2PACX-1vTyb7i8hKbLErLrmPmiRSnOIhsinOaYXw3YX4pZnpFB_a_Fow-QaK19GmcoAvneYolwJDTHtLBEwybm/pub?gid=1899213608&single=true&output=csv')

# Wide to long

df <- df |>
  pivot_longer(
    cols = starts_with("Book"), # Adjust this to match your column names
    names_to = c("Book", "Round"),
    names_pattern = "Book (\\d+), Round (\\d+)", # Regex to extract Book and Guess numbers
    values_to = "Guess" # Name of the column for the values
  )

# Change Book 1 to Decentralized and Book 2 to Centralized
df <- df |>
  mutate(Network_Type = ifelse(Book == 1, "Decentralized", "Centralized")) |>
  select(-Book) |>
  mutate(Guess = as.numeric(Guess)) |>
  mutate(Round = as.numeric(Round)) |>
  filter(!is.na(Guess)) # Remove rows with NA values in 
```
```{r}
# Filter to timestamps within the last 30 days
df <- df |>
  mutate(timestamp = as.POSIXct(Timestamp, format = "%m/%d/%Y %H:%M:%S")) |>
  filter(timestamp >= Sys.time() - 30 * 24 * 60 * 60) # Last 30 days

```

## Today's Dad Joke

When I was young, we were really poor. My family could only afford a second-hand calculator, which was missing the "X" button.

<p class='fragment'>
Times were hard.
</p>

## Housekeeping

> - Exam next week
> - One page of notes, front and back (printed)
>	- Short response + multiple choice
> - Making networks without grids
>   -  `set_graph_style(family='sans')` or `theme_graph()`

<!--
## AI Conversation

> - There is a tension
>   - Universities are meant to educate and credential
>   - AI is an amazing tutor
>   - But it can also be used to substitute for learning and make evaluation much more difficult
>   - AI can help us to understand deeper, or to pretend that we understand deeper.

## AI Conversation

> - Imagine that you were the instructor for a class like COM 411. How would you design the class to encourage learning given today's technological environment?
-->

## Final Project

> - Timing for [final project](../../assignments/final_project/Final-project.qmd)
>	- Don't wait - get started now!
>	- Reach out if you get stuck
> - Look at example projects from the past
> - Not all visualizations have to be networks
> - Concepts + Visualizations

# Mini-lecture on Wisdom of Crowds and Collective Intelligence

## Collective Intelligence

> - Idea that groups can be smarter than the smartest person in the group
> - Sometimes it's simple; e.g, aggregating information from multiple sources
> - Sometimes there is an emergent property, where the group thinks and behaves differently

## Ants

{{< video https://www.youtube.com/watch?v=IQW_K8rq1N4 height="85%" width="100%" >}}

## Organizations

> - While organizations are much more hierarchical than ants, they can still have emergent properties
> - Partially a function of history, partially of the structure of the organization
> - Even a culture everyone dislikes can be hard to change
> - E.g., metric system, QWERTY keyboards

## Cities

> - High-level decisions about zoning, infrastructure, etc. but the character of a city is emergent
> - Jane Jacobs: "The Death and Life of Great American Cities"


## Wisdom of Crowds Activity

> - Get out a piece of paper
> - Guess how many pages are in this book and write down your guess
> - Tell your guess (and no other information) to the two people on either side of you
> - Guess again how many pages are in the book and write it on a new line (Guess 2)

## Activity 2

> - Guess 3: Guess the pages in this book
> - Two random people will tell us their guess
> - Guess 4: Guess the pages again for this book


## Collective understanding and decision-making

> - Wisdom of crowds is idea that in many contexts, the average is a really good guess!
> - What is groupthink? Have you seen examples? How might this relate?
> - Dr. Josh Becker et al. show that in some cases, communication can help
>   - Those whose guesses are worst move the most
>   - Divorced from other social relationships (not multiplex)


## Results

## Decentralized network

```{r}
df |>
  filter(Network_Type == 'Decentralized') |>
  ggplot() +
  geom_density(aes(x = Guess, fill = as_factor(Round)), alpha = .4) + #, position='dodge', bins = 5) + 
  geom_vline(xintercept=BOOK_1_LENGTH, color = 'red', linetype='dashed') + 
  theme_minimal() + 
  scale_fill_viridis_d(name='Round', end=.8)
```

## Centralized network

```{r}
df |>
  filter(Network_Type == 'Centralized') |>
  ggplot() +
  geom_density(aes(x = Guess, fill = as_factor(Round)), alpha = .4) + #, position='dodge', bins = 5) + 
  geom_vline(xintercept=BOOK_2_LENGTH, color = 'red', linetype='dashed') + 
  theme_minimal() + 
  scale_fill_viridis_d(name='Round', end = .8)
```

## Connections to other network concepts

> - How does this relate to our week on small group networks?
>   - Transactive memory
>   - Hierarchy
>   - Centralization

## Connections to other network concepts
> - How could you design a group / network to get the benefits of the wisdom of crowds?
> - When would you expect a wisdom of crowds approach to work well?
> - When would you expect it to fail?

## Discussion Questions

> - What are the most practical ways to design communication networks or platforms that promote decentralized influence?
> - What are some real-life examples of collective intelligence making smart decisions regarding norms, such as when IP implicitly became the norm for network communication, as mentioned in the discussion video?
> - If collective intelligence relies on social structures more than on individual skills, then what might schools, organizations, or communities do to foster environments that promote informed participation and the real "wisdom of the crowd"?

<aside class="notes">
- If there's time, talk about R
</aside>

## Plan for Today

> - Exam Review


## Exam Review

>	- Visualization / ggraph
>   - Look at code and explain what it will look like
>   - Find the bug in code

## Example

- What code produces this graph?


```{r, echo=FALSE}
G <- play_gnp(18, .15, directed = F) |> as_tbl_graph()
set_graph_style(family='sans')

G |>
  mutate(centrality = centrality_degree()) |>
  activate(edges) |>
  ggraph() +
  geom_edge_fan(color = 'purple') +
  geom_node_point(aes(size = centrality), color = 'green')
```


```r
G |>
  mutate(centrality = centrality_degree()) |>
  activate(edges) |>
  ggraph() +
  geom_edge_fan(color = 'purple') +
  geom_node_point(aes(size = centrality), color = 'green')
```


## Exam Review

What would this code do?

```r
G |>
  activate(nodes) |>
  mutate(ec = centrality_eigen()) |>
  filter(ec > 1) |>
  ggraph() +
  geom_edge_fan(color = 'orange') +
  geom_node_point(aes(size = ec), color = 'green')

```

## Exam Review

> - Centrality
> - Network representations (Edgelists and matrices)
> - Network formation
>   - Homophily, focus theory, transitivity, social exchange theory, etc.
> - Tie strength and small worlds

## Exam Review

> - Power and social capital
> - Contagion and diffusion
> - Centrality
> - Bipartite networks and projections
> - Random rewirings

## Exam Review

> - Friendship paradox
> - Strength of weak ties
> - Wisdom of crowds / collective intelligence

<!--
## Review Questions

> - What is the difference between a filter bubble and an echo chamber?
> - How do echo chambers relate to social influence?
> - What did Chris Bail et al.'s research find about breaking out of echo chambers?
> - How has the internet changed social networks? What are the most important positives and negatives?


## Review Questions

> - What is context collapse?
> - How does context collapse relate to network principles we have discussed?

## Discussion Questions
-->