---
title: Week 6 - Centrality and Centralization
sort: 6
format: 
  revealjs:
    css: /resources/custom_css.css
    theme: white
    progress: true
    transition: convex
    reveal_plugins: ['notes']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = './')
source("resources/preamble.R")
library(tidygraph)
library(ggraph)
```

## Dad Joke

Can you believe my wife told me to quit acting like a flamingo?

<p class='fragment'>
I had to put my foot down.
</p>

## Housekeeping

> - Self Assessment Reflection Feedback coming soonish
> - Homework questions on Discord
>	- How to copy and paste code
> - A lot like qmd files

<div class='fragment'>
<code><pre>

```r

G |>
  activate(nodes) |>
  filter(age > 25)
```

</pre>
</code>
</div>

## Housekeeping

> - How did the AI assignment go?
>   - What was it good at?
>   - What was it bad at?


<!--
## Collective Norms Discussion

> - I think our in-person conversations could be better
> - Too many tech distractions
> - Coming prepared
> - How to have more natural conversations?
-->


## Concept Review

> - What is centrality?
> - What is the difference between centrality and centralization?

## Flash presentations

> - Choose one centrality measure
> - 4 minutes of prep
> - 1-2 minutes of presentation
> - What is it? How is it calculated? What does it tell us about power/influence?

<!--
## Concept Review

> - What is betweenness centrality?
> - What is closeness centrality?
> - What is eigenvector centrality?
-->

## Concept Review

> - What aspects of power/influence do centrality measures miss?
> - What can centralization tell us about a network?
> - What does density tell us about a network?
> - What is hierarchy?

## Practical Review

What is the degree/closeness/betweenness centrality of node A? What is the density of the network?

```{r echo=FALSE}
set.seed(123)
G <- play_gnp(6, 0.5, directed=F) |> as_tbl_graph() |>
  mutate(name = LETTERS[1:nrow(.N())])


G |>
  # Get first N letters of alphabet
  ggraph(layout = 'stress') +
  geom_edge_fan() +
  #geom_node_point(aes(size = centrality_degree())) +
  geom_node_label(aes(label = name), size = 8) +
  theme_graph()
```

## Practical Review

What is the degree/closeness/betweenness centrality of node A?

```{r}
G |>
  mutate(degree = centrality_degree(),
         closeness = centrality_closeness(),
         betweenness = centrality_betweenness()) |>
  filter(name == "A")

```


## Discussion Questions

> - Could someone have high network centrality but low actual power/influence?
> - Does network position cause influence, or vice versa?
> - Bipartite graphs (Paul Revere)
> - How do these approaches scale to huge networks?
> - How are centrality measures impacted by missing data?


## R Review

